### 2021-04-15 Paper/Main.tex
why not another push

\begin{document}
\section{Introduction and Motivation}
Asset pricing has a long history of models and theories governing how investors make decisions. Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns. Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework. ESG investing considers how firms respond to these factors as potential impetuses for excess performance. This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape. Morgan Standley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}. Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.

@@

\section{Data}
\label{Data}
The data used for this paper was sourced from three datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily average returns for the MSCI World ESG Focus Index and S/&P petroleum index, respectively. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.  An ESG factor, denoted as EMP (ESG minus petroleum), was created by taking the  MSCI World ESG Focus Index returns and subtracting them from the  S\&P petroleum index. Petroleum companies act as a stand-in for the opposite of ESG as the petroleum space is particularly inept at adapting to ESG  changes, a fact well documented in academic literature \cite{Frynas2005TheFD}. In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals & Mining Select Industry Index as the contra ESG portfolio for the EMP factor. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.

@@

In the preliminary analysis of the ESG factor I sort securties  on beta as well as size deciles. Each security will be assigned a decile for beta and size according the the CRSP /NYSE american beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size. After sorting each of the securities by Beta and size decile  300,264 of the original 830,889 data points had matches for both portfolios. To estimate the factor sensitivities base on the discribed decile groups I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.

\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA_t+e_iEMP_t
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}

@@

\section{Results}
\label{Results}
Table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}. The column axis is the distinct size deciles where 1 is the smallest market cap companies and 10 is the declie with the largest market cap companies. The row axis includes estimated coefficients for the intercept, five Fama French factors and the EMS factor. $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor. In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts. Moving onto the second step in the two pass sort to asess the risk premia,

Moving onto the analysing the results of the second stage in the two pass sort we would expect to see that   
% size cofficents
\begin{center}
\input{Paper/prelim_size}\\
\textbf{Table 3:} Size decile by estimated coefficients for the  five fama french factors and EMP factor
\end{center}
% size p vals
\begin{center}
\input{Paper/p_vals_size_prelim}\\
\textbf{Table 4:} Size decile by p-values for table 3
\end{center}
% beta cofficents
\begin{center}
\input{Paper/prelim_beta}\\
\textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
\end{center}
% beta p vals
\begin{center}
\input{Paper/p_vals_beta_prelim}\\
\textbf{Table 6:} Size decile by p-values for table 5
\end{center}

>>>>

\begin{document}
\section{Introduction and Motivation}
Asset pricing has a long history of models and theories governing how investors make decisions. Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns. Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework. ESG investing considers how firms respond to these factors as potential impetuses for excess performance. This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape. Morgan Stanley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}. Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.

@@

\section{Data}
\label{Data}
The data used for this paper was sourced from three datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.

An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index. Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to the what Fama french do to isolate the specific risk component. The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis. In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}. Although the Petroleum industry has major environmental, risks there is less social orientated risks as the index \$VICE which is an index of alcohol, drug, gaming and betting industry might provide a better exposure to the broad based effects of social ESG risk factors as a contra ESG portfolio. However \$Vice has only traded for two years, as such the data is far less robust. Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the  In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.

\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

@@

For this analysis I sort each security by month into decile groups for beta as well as size (based on market cap). Each security will be assigned a decile for beta and size according the the CRSP /NYSE american beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size. After sorting each of the securities by Beta and size decile  300,264 of the original 830,889 data points had matches for both portfolios. To estimate the factor sensitivities base on the discribed decile groups I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.

\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}

@@

\section{Results}
\label{Results}
Table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}. The column axis is the distinct size deciles where 1 is the smallest market cap companies and 10 is the declie with the largest market cap companies. The row axis includes estimated coefficients for the intercept, five Fama French factors and the EMS factor. $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor. In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts. Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3. The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level. It is important to note that all of the factors are less significant that largest and small polios indicating that the data might have point at far ends which do not behave like the other deciles.  Moving to table 6 and the second step in the two pass sort we can see the risk premia for each of the factors. All of the Fama French factors in the size decile two pass sort have positive risk premia. The EMP factor large than all of the other factors with a coefficient of approximately  16.8\%

% size cofficents
\begin{center}
\input{Paper/prelim_size}\\
\textbf{Table 3:} Size decile by estimated coefficients for the  five fama french factors and EMP factor
\end{center}
% size p vals
\begin{center}
\input{Paper/p_vals_size_prelim}\\
\textbf{Table 4:} Size decile by p-values for table 3
\end{center}
% beta cofficents

Moving on to the beta decile based two pass sort the reuslts are quite similar to the size decile results. Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile where 1 is the portfolio comprised of the companies with lowest bet and 10 is the portfolio of companies  with the greatest beta values.  Similar to the size decile the EMP factor exhabits a negative relations with regards to its sensitivity. However, unlike the size decile analyis there does not appear to be a signficant relationship between the  beta and the sensitivity of the factor. Although the largest impact is on the lowest decile the least impact is experienced on the 6th decile porfolio and the largest deciel is close to the middle in terms of sensitivity. Similar to the size decile analysis the EMP coefficients are all very significant with every decile having significance at the 5\% level. When considering the second phase of the two pass sort the similarities to the size decile continue where in each of the Fama French five factors exhabit a positive risk premia and the EMP also exhabits a positive relationship. the EMP factor again has the high coefficient with approximately 15.8\%.

\begin{center}
\input{Paper/prelim_beta}\\
\textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
\end{center}
% beta p vals
\begin{center}
\input{Paper/p_vals_beta_prelim}\\
\textbf{Table 6:} Size decile by p-values for table 5
\end{center}
% pass 2
\begin{center}
\input{Paper/pass_2}\\
\textbf{Table 7:} second pass results for the size and beta decile portfolios
\end{center}
